---
title: "The KnapsackPack package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{The KnapsackPack package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(KnapsackPack)

# Example data
suppressWarnings(RNGversion(min(as.character(getRversion()),"3.5.3")))
set.seed(42, kind = "Mersenne-Twister", normal.kind = "Inversion")
n <- 1e6
knapsack_objects <-
data.frame(
w=sample(1:4000, size = n, replace = TRUE),
v=runif(n = n, 0, 10000)
)

```

## Brute force
```{r}
# Runtime for 16 objects
system.time(brute_force_knapsack(x = knapsack_objects[1:16,], W = 2000))
```

## Dynamic programming
```{r}
# Runtime for 500 objects
system.time(knapsack_dynamic(x = knapsack_objects[1:500,], W = 2000))
```

## Greedy heuristic
```{r}
# Runtime for 1 000 000 objects
system.time(greedy_knapsack(x = knapsack_objects, W = 2000))
```

## Code profiling and optimization
By profiling our code we managed to cut the runtime by about 30%. For example, we indexed vectors instead of data frame (as we initially did), that seemed to have improved the efficiency.

## Parallel brute force search
```{r}
# Non parallel
system.time(brute_force_knapsack(x = knapsack_objects[1:20,], W = 2000, parallel=FALSE))

# Parallel
system.time(brute_force_knapsack(x = knapsack_objects[1:20,], W = 2000, parallel=TRUE))
```

The parallelized brute search has multiple cores and uses the *apply* function which is expected to be faster than the one core and for loop brute search method, however, it rendered a longer runtime. The reason for this could be due to the simplicity of the calculations in the for loop which makes it faster.


